Galois Slicing as Differentiation

### 2025-07-11 main/related-work.tex
[ tex ] tweak related work

\emph{Tangent categories}, due originally to \citet{rosický84} and developed by \citet{cockett14,cockett18},
provide an abstract categorical framework for reasoning about differentiation, inspired by the structure of
the tangent bundle in differential geometry. In a tangent category, each object $X$ is equipped with a tangent
bundle $T(X)$, and each morphism $f: X \to Y$ has a corresponding differential map $T(f): T(X) \to T(Y)$
satisfying axioms analogous to the chain rule and linearity of differentiation. Tangent categories generalise
Cartesian differential categories \cite{cdcs}, which model differentiation over cartesian closed categories
using a syntactic derivative operator. Reverse Tangent categories \cite{reverse-tangents} further axiomatise
the existence of reverse derivatives. In \conref{tangent-stable-fns} and \conref{tangent-fam}, we have
conjectured that the categories we have identified in this paper as models of \GPS are Tangent categories. As
well as providing some substantiation to the claim that \GPS is a form of differentiation, we should be able
to apply general results about Tangent categories to our setting. % BOB: such as?

>>>>

\emph{Tangent categories}, due originally to \citet{rosický84} and developed by \citet{cockett14,cockett18},
provide an abstract categorical framework for reasoning about differentiation, inspired by the structure of
the tangent bundle in differential geometry. In a tangent category, each object $X$ is equipped with a tangent
bundle $T(X)$, and each morphism $f: X \to Y$ has a corresponding differential map $T(f): T(X) \to T(Y)$
satisfying axioms analogous to the chain rule and linearity of differentiation. Tangent categories generalise
Cartesian differential categories \cite{cdcs}, which model differentiation over cartesian closed categories
using a syntactic derivative operator. Reverse Tangent categories \cite{reverse-tangents} further axiomatise
the existence of reverse derivatives. In \conref{tangent-stable-fns} and \conref{tangent-fam}, we have
conjectured that the categories we have identified in this paper as models of \GPS are Tangent categories. As
well as providing some substantiation to the claim that \GPS is a form of differentiation, this would also
clarify the role of higher derivatives in \GPS, which we conjecture are related to \emph{program
  differencing}.
